### 2021-04-22 newvariant.tex
Handing back

\begin{figure}[!t]
\includegraphics[width=\textwidth]{relstrength.pdf}
\caption{
\textbf{Relative strength of the new variant conditional on observed relative speed under five epidemiological conditions.}
The relative strength of the new variant $\rho$ conditional on the speed of the wild type $\rw$, the ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and the squared coefficient of variation in generation intervals $\kappa$.
The relative strength of the new variant $\rho$ is calculated using $\delta=0.1\pday$, $\Gw = 5\days$, and $\kappa = 1/5$.
\jd{If you add $\rw$ to the figure titles, you could skip this sentence and make life easier for readers.} Five epidemiological conditions assume $\rw=-0.15\pday$ ($\rw < \rv < 0$), $\rw=-0.1\pday$ ($\rw < \rv = 0$), $\rw=-0.05\pday$ ($\rw < 0 < \rv$), $\rw=0\pday$ ($0 = \rw < \rv$), and $\rw=0.05\pday$ ($0 < \rw < \rv$).
}
\label{fig:relstrength}
\end{figure}

@@

Based on the observed relative growth rate of B.1.1.7 in the UK, we start by  taking the relative speed of the variant to be $\delta = 0.1/\textrm{day}$  \citep{davies2021estimated}; the mean generation interval of the wild type to be $\Gw = 5\days$ \citep{ferretti2020quantifying}; and the squared coefficient of variation of generation intervals to be $\kappa=0.2$ \citep{ferretti2020quantifying}.
We consider the spread of B.1.1.7 as an example but our qualitative conclusions should hold for other VoCs.
As noted above, we assume that the variant and the wild type have equal $\kappa$ throughout, and only consider differences in the mean.
We evaluate the estimates of relative strength $\rho$ across a wide range of $\kappa$ from 0 (fixed-length generation intervals) to 1 (exponential distribution).
To further explore how inference depends on underlying epidemiological conditions (i.e., whether the incidence of the variant and the wild type are growing or shrinking), we consider five epidemiological conditions: (1) $\rw < \rv < 0$, (2) $\rw < \rv = 0$, (3) $\rw < 0 < \rv$, (4) $0 = \rw < \rv$, and (5) $0 < \rw < \rv$. \jd{This looks like the right order to me; it does not seem to match F1.}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{relspeed.pdf}
\caption{
\textbf{Relative speed of the new variant given assumed relative strength under five epidemiological conditions.}
The relative speed of the new variant $\delta$ given strength of the wild type $\Rw$, ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and squared coefficient of variation in generation intervals $\kappa$.
Relative speed of the new variant $\delta$ is calculated using $\rho=1.61$, $\Gw = 5\days$, and $\kappa = 1/5$.
Five epidemiological conditions assume $\Rw=1/\rho^{3/2}$ ($\Rw < \Rv < 1$), $\Rw=1/\rho$ ($\Rw < \Rv = 1$), $\Rw=1/\rho^{1/2}$ ($\Rw < 1 < \Rv$), $\Rw=1$ ($1 = \Rw < \Rv$), and $\Rw=\rho^{1/2}$ ($0 < \Rw < \Rv$).
}
\label{fig:relspeed}
\end{figure}

@@

Here, we investigate how misspecification of the generation-interval distribution of the variant affects our inference of relative strength from inference data under the assumption that the true generation-interval distribution of the wild type is known.
We use a two-strain renewal equation that assumes perfect cross-immunity to simulate three different scenarios (see Methods):  
(1) the wild type and the variant have the same generation-interval distributions (which match the known distribution, shown in \fref{Rtbias}A--C);
(2) the variant has a shorter mean generation interval (\fref{Rtbias}D--F); and
(3) the variant has a longer mean generation interval (\fref{Rtbias}G--I).
Then, we compare the estimated ratio $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ with the true ratio $\rho = \Rv(t)/\Rw(t)$.
In order to allow for changes in $\RR$ driven by the introduction and lifting of non-pharmaceutical interventions \citep{flaxman2020Rt}, we model $\Rw(t)$ using a step function (\fref{Rtbias}) and let $\Rv(t) = \rho \Rw(t)$.

@@

When the assumed distribution matches the true distribution, the estimated reproduction numbers match the true values (\fref{Rtbias}A), and therefore the ratio remains constant (\fref{Rtbias}B).
However, if the generation-interval distribution of the variant differs from the assumed distribution, the ratio changes over time (\fref{Rtbias}E,H).
In particular, if the true generation intervals have a shorter mean than the assumed distribution, we over-estimate $\Rv(t)$ during the growth phase and under-estimate during the decay phase (and vice versa, \fref{Rtbias}D,G), which further translates to biases in the estimated relative strength (\fref{Rtbias}E,H).
Therefore, biases in the estimated relative strength will correlate with the strength of the variant, (i.e., whether $\Rv(t)$ is below or above 1).

In practice, estimates of instantaneous reproduction numbers $\RR(t)$ (and therefore, their ratios) can be noisy due to limited data availability or model assumptions;
instead, we might want to estimate a single value of relative strength $\rho$.
For example, we can estimate $\rho$ by plotting the estimated strength of the variant $\hat{\RR}_{\textrm{var}}(t)$ against the estimated strength of the wild type $\hat{\RR}_{\textrm{wt}}(t)$---as presented in Figure 2 of \cite{volz2021transmission}---and performing a linear regression (\fref{Rtbias}C,F,I).
When generation-interval distributions are correctly specified, we obtain a straight line with a slope of $\rho$ and intercept at zero (\fref{Rtbias}C).
However, when the assumed mean generation interval is longer than the that of the variant, we over-estimate the slope (and vice versa, \fref{Rtbias}F,I).
In theory, we expect the regression line to go through the origin, but when the assumed generation-interval distribution differs from the true distribution, this is not necessarily the case: biases in slopes due to mis-specified generation-interval distributions also lead to biases in intercepts.
This analysis suggests another method for assessing potential biases in the estimates of relative strength.

@@

We consider two types of intervention:
an intervention of constant strength, which reduces transmission by a constant factor $\theta$ regardless of age of infection ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$); and an intervention of constant speed, which reduces transmission after infection by a constant rate $\phi$ ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$).
In this case, we can control the spread of the variant when $\theta > \Rv$ or $\phi > \rv$, respectively.
As a baseline scenario, we consider constant-strength and speed interventions that reduce $\Rv$ to 0.9 when both the variant and the wild type have equal mean generation intervals (\fref{strengthspeed}A--B).
While both interventions are equally effective on strength-scale (that is, $\Ry{post}=\int  K_{\mathrm{post}}(\tau) \dtau = 0.9$), they have different dynamical implications.
The constant-strength intervention affects transmission equally throughout the course of infection, whereas the constant-speed intervention has greater impact of transmission that occur later;
as a result, the constant-speed intervention reduces the post-intervention mean generation interval (\fref{strengthspeed}B) and leads to (slightly) faster exponential decay (therefore, lower $\ry{post}$).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{control.pdf}
\caption{
\textbf{Effects of constant-strength and speed of interventions against the spread new variant.}
(A--B) Pre-intervention (black) and post-intervention kernel (colored) of the new variant under constant-strength (A) and speed (B) interventions assuming that the variant has equal generation intervals as the wild type.
(C--D) Pre-intervention (black) and post-intervention kernel (colored) of the new variant under constant-strength (C) and speed (D) interventions assuming that the variant has longer generation intervals as the wild type.
(E) Pre-intervention (black) and post-intervention strength conditional on the mean generation interval of the variant (colored).
(F) Pre-intervention (black) and post-intervention speed conditional on the mean generation interval of the variant (colored).
Strength and speed are calculated assuming$\rw=0\,\textrm{/days}$, $\delta=0.1\,\textrm{/days}$, $\rv=0\,\textrm{/days}$, $\Gy{wt}=5\,\textrm{days}$, and $\kappa=1/5$ for pre-intervention conditions.
Constant strength and speed interventions are modeled such that post intervention strength of the new variant, assuming equal mean generation intervals at 5 days, corresponds to 0.9.
}
\label{fig:strengthspeed}
\end{figure}

@@

The constant post-intervention speed can be understood by considering a speed-based decomposition of the kernel: $K_{\mathrm{pre}}(\tau) = \exp(\rv) b(\tau)$, where $b(\tau)$ is the initial backward generation-interval distribution and therefore integrates to 1 \citep{doi:10.1098/rspb.2020.1556}.
This decomposition allows us to re-write the post-intervention kernel (under the constant-speed intervention) as follows:
\begin{equation}
K_{\mathrm{post}}(\tau) = \exp(\rv-\phi) b(\tau).
\end{equation}
Since the post-intervention speed $\ry{post}$ is determined by the following implicit equation,
\begin{equation}
1 = \int_0^\infty \exp(-\ry{post} \tau) K_{\mathrm{post}}(\tau) \dtau = \int_0^\infty \exp(\rv-\phi-\ry{post}) b(\tau)  \dtau,
\end{equation}
we have $\ry{post} = \phi - \rv$. Therefore, if we can reduce infectiousness at a faster rate than the observed speed of spread (i.e., if the speed of intervention is faster than the epidemic speed, $\phi > \rv$), we can control the epidemic regardless of the underlying generation-interval distribution.

@@

We explored how the generation-interval distribution shapes the link between relative strength and speed of the new variant.
Longer generation intervals generally lead to higher relative strength for a given relative speed (and conversely, lower relative speed for a given relative strength); these can be reversed when incidence is decreasing.
Neglecting potential differences in the mean generation intervals between the variant and the wild type can bias estimates of the relative strength from incidence data;
these biases may be systematically assessed by considering whether apparent changes in relative strength over time are associated with underlying epidemiological dynamics of the variant and the wild type.
Finally, differences in generation intervals can further lead to different conclusions about the effectiveness of interventions:
if the variant has longer (respectively, shorter) generation intervals than the wild type, speed-like (strength-like) interventions, will be more relatively more effective than naive estimates would suggest.

As the new variants, in particular the B.1.1.7 variant, of SARS-CoV-2 are spreading and becoming dominant in many countries, it is clear that the variants are more transmissible (have higher strength) than the wild type \citep{switzerland2021variant, davies2021estimated, di2021impact, leung2021early, volz2021transmission,zhao2021}.
Our analysis supports estimates of a higher strength of the new variant bounding on a wide range of parameter ranges about epidemiological conditions and generation-interval distributions.
However, our analysis also shows that uncertainty in generation-interval distributions must be taken into account to obtain accurate estimates of the relative strength of variants.
If the variant has a longer infectious period \citep{kissler2021densely}, and therefore extended generation-interval distributions, current estimates may be underestimating the true relative strength.

There is currently a (minor) discrepancy in the estimates of relative strength of new variants, particularly for B.1.1.7.
Mathematical analyses have typically reported greater than 1.4 fold increase in reproduction number for the B.1.1.7 variant ($\rho > 1.4$) whereas an independent analysis of secondary attack rates from contact tracing data suggests 1.25--1.4 fold increase \citep{ukinvest}.
We can ask whether different estimates can be reconciled by generation intervals alone---for simplicity, we assume that the variant and the wild type have identical generation-interval distributions as other studies have.
For example, \cite{davies2021estimated} estimated $\rho=1.77$ assuming a fixed generation generation interval at 5.5 days from $\delta=0.104\pday$ in their regression analysis;
similarly, \cite{volz2021transmission} estimated $\rho=1.79$ by directly comparing the ratio between $\Rv$ and $\Rw$ assuming a mean generation interval of 6.5 days and a squared coefficient of variation of 0.38.
While \cite{volz2021transmission} assumed a longer mean generation interval (which typically increases $\rho$), their estimate is similar to \cite{davies2021estimated}, likely because the effect of a wide distribution (which typically reduces $\rho$) can cancel out with the effect of longer intervals---in fact, we obtain $\rho=1.81$ from $\delta=0.104\pday$, $\rw=0\pday$, $\Gv=\Gw=6.4\days$, and $\kappa=0.38$ under the gamma assumption, explaining the lack of differences in the estimates of $\rho$ despite the differences in the assumed generation-interval distributions.

However, a fixed generation generation interval assumed by \cite{davies2021estimated} is too narrow.
The mean generation interval of 6.5 days assumed by \cite{volz2021transmission} is also likely too long:
first, this estimate actually corresponds to the observed mean serial intervals from China in between January and February, 2020 \citep{bi2020epidemiology}, during which they were more likely to observe longer serial intervals due to initial epidemic growth \citep{park2021forward, ali2020serial}; and second, recently infected individuals may be less likely to transmit after symptom onset due to increased awareness of the pandemic compared to earlier on, leading to shorter generation intervals.
Instead, if we replace the assumed generation-interval distributions for \cite{volz2021transmission} with shorter generation interval estimates from Tianjin, China ($\Gw=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.27$, which is consistent with the attack rate analysis \citep{ukinvest}.
We do not claim that these estimates of $\rho$ or $\Gw$ are more accurate;
instead, this back-of-the-envelope calculation highlights the importance of propagating uncertainty in the generation-interval distribution in assessing the relative strength and speed of SARS-CoV-2 variants.
Obtaining an accurate and precise estimate of $\rho$ will need to take into account changes in $\rho$ as well as underlying $\rw$.

@@

We also assumed that the incidence of infection caused by the variant and the wild type is known.
However, estimating $\RR(t)$ from real data, such as case incidence, is often associated with several practical challenges, such as delays between infection and reporting and changes in testing patterns \citep{gostic2020practical}.
We have chosen to focus here on the underlying dynamical mechanisms that may impact inference.
Future studies may consider the development of methods based on mechanism outlined here to assess potential biases in their estimates of relative strength.

In moving from mechanism to improved estimates, we argue that both perspectives (i.e., whether we hold relative strength or speed constant) are useful in understanding the dynamics of the new variant.
Early in the spread of an emerging variant, it is likely more convenient to fix the relative speed, given that speed can be directly observed.
As more information about the transmission and immunity profiles of the new variant becomes available, we advise fixing the relative strength and inferring the speed as this assumption better matches biological mechanisms for its higher strength.
Researchers should be mindful about which quantity they hold constant and how their assumptions lead to their conclusions \citep{doi:10.1098/rspb.2020.1556}.

Here, we lay out a few practical advice in analyzing the epidemiological dynamics of new variants for future studies.
First, models that assume a constant relative speed, such as standard logistic growth model, should be avoided for characterizing the growth of new variants---the relative speed changes over time depending on epidemiological conditions.
Second, the absolute strength and speed should not be neglected over their relative values, and both relative and absolute values should be reported together.
While the relative strength and speed are useful for describing the spread of new variants, the absolute values determine their spread and control.
In addition, as we showed in our analysis, it is critical to understand the underlying epidemiological conditions of the variant and the wild type in interpreting the relative strength and speed.
Finally, uncertainty in generation intervals should be carefully considered.

Even though SARS-CoV-2 has been spreading for more than a year, there is still considerable uncertainty in its generation intervals.
Observational studies typically focus on serial intervals (i.e., time between symptom onset of the infector and the infectee; \cite{svensson2007note}) because they are easier to measure \citep{griffin2020rapid}.
Furthermore, realized serial intervals are subject to dynamical biases that can be difficult to tease apart \citep{park2021forward}.
A few studies have tried to estimate the generation-interval distribution from serial intervals, with means ranging between 3--6 days and squared coefficients of variation ranging between 0.1--1 \citep{ferretti2020quantifying,Ferretti2020timing,ganyani2020estimating,knight2020estimating}; 
as we have shown, this uncertainty alone can generate large uncertainty in the inference of relative strength.
Furthermore, most estimates neglect dynamical biases in serial intervals caused by underlying epidemiological dynamics in their estimates of generation intervals.
Serial intervals also do not account for asymptomatic transmission, adding further uncertainty to inferences of speed and strength \citep{park2020time}.

Since the beginning of the SARS-CoV-2 pandemic, a few studies have emphasized the relevance of leveraging generation interval distributions to improve estimates of strength (e.g., \cite{doi:10.1098/rsif.2020.0144,park2021forward}) and the relative importance of modes of transmission (e.g., asymptomatic vs. symptomatic, \cite{park2020time}).
These same lessons should be considered when assessing the spread of variants in a partially susceptible population and improving efforts to control spread.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions.

The spread of new SARS-CoV-2 variants and the replacement of previously dominant lineages represent ongoing challenges for controlling the SARS-CoV-2 pandemic.  
By explicitly considering epidemiological context and generation interval differences together, we have shown that improving estimates of the the relative duration of infectiousness at the individual scale may represent a pathway towards more effective interventions. 
Specifically, we have shown that speed-like interventions, such as contact tracing, will be more effective if variants have longer generation intervals.
Most intervention strategies throughout the current pandemic have focused on strength-like interventions \citep{flaxman2020Rt}, such as lock-downs, partly because pre-symptomatic transmission of SARS-CoV2 have limited the effectiveness of contact tracing efforts \citep{hellewell2020feasibility}.
However, given the possibility that new variants can have different infection characteristics, future studies should consider whether their transmission dynamics also differ (e.g., the amount of pre-symptomatic transmission) and evaluate intervention strategies accordingly.

>>>>

\begin{figure}[!t]
\includegraphics[width=\textwidth]{relstrength.pdf}
\caption{
\textbf{Relative strength of the new variant conditional on observed relative speed under five epidemiological conditions.}
The relative strength of the new variant $\rho$ conditional on the speed of the wild type $\rw$, the ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and the squared coefficient of variation in generation intervals $\kappa$.
The relative strength of the new variant $\rho$ is calculated using $\delta=0.1\pday$, $\Gw = 5\days$, and $\kappa = 1/5$.
\jd{If you add $\rw$ to the figure titles, you could skip this sentence and make life easier for readers.} Five epidemiological conditions assume $\rw=-0.15\pday$ ($\rw < \rv < 0$), $\rw=-0.1\pday$ ($\rw < \rv = 0$), $\rw=-0.05\pday$ ($\rw < 0 < \rv$), $\rw=0$ ($0 = \rw < \rv$), and $\rw=0.05\pday$ ($0 < \rw < \rv$).
}
\label{fig:relstrength}
\end{figure}

@@

Based on the observed relative growth rate of B.1.1.7 in the UK, we start by  taking the relative speed of the variant to be $\delta = 0.1/\pday$  \citep{davies2021estimated}; the mean generation interval of the wild type to be $\Gw = 5\days$ \citep{ferretti2020quantifying}; and the squared coefficient of variation of generation intervals to be $\kappa=0.2$ \citep{ferretti2020quantifying}.
We consider the spread of B.1.1.7 as an example but our qualitative conclusions should hold for other VoCs.
As noted above, we assume that the variant and the wild type have equal $\kappa$ throughout, and only consider differences in the mean.
We evaluate the estimates of relative strength $\rho$ across a wide range of $\kappa$ from 0 (fixed-length generation intervals) to 1 (exponential distribution).
To further explore how inference depends on underlying epidemiological conditions (i.e., whether the incidence of the variant and the wild type are growing or shrinking), we consider five epidemiological conditions: (1) $\rw < \rv < 0$, (2) $\rw < \rv = 0$, (3) $\rw < 0 < \rv$, (4) $0 = \rw < \rv$, and (5) $0 < \rw < \rv$. \jd{This looks like the right order to me; it does not seem to match F1.}

@@

\begin{figure}[!th]
\includegraphics[width=\textwidth]{relspeed.pdf}
\caption{
\textbf{Relative speed of the new variant given assumed relative strength under five epidemiological conditions.}
\jd{Again, I'd like to see these ordered better. I also still like the idea of putting the value of Rwt in the figure titles (the geometric progression is cute, and we should keep those values, but maybe just mention it briefly and show the values to 2dp.} The relative speed of the new variant $\delta$ given strength of the wild type $\Rw$, ratio between the mean generation interval of the new variant $\Gv$ and that of the wild type $\Gw$, and squared coefficient of variation in generation intervals $\kappa$.
Relative speed of the new variant $\delta$ is calculated using $\rho=1.61$, $\Gw = 5\days$, and $\kappa = 1/5$.
Five epidemiological conditions assume $\Rw=1/\rho^{3/2}$ ($\Rw < \Rv < 1$), $\Rw=1/\rho$ ($\Rw < \Rv = 1$), $\Rw=1/\rho^{1/2}$ ($\Rw < 1 < \Rv$), $\Rw=1$ ($1 = \Rw < \Rv$), and $\Rw=\rho^{1/2}$ ($0 < \Rw < \Rv$).
}
\label{fig:relspeed}
\end{figure}

@@

Here, we investigate how misspecification of the generation-interval distribution of the variant affects our inference of relative strength from inference data under the assumption that the true generation-interval distribution of the wild type is known.
We use a two-strain renewal equation that assumes perfect cross-immunity to simulate three different scenarios (see Methods):  
(1) the wild type and the variant have the same generation-interval distributions (which match the known distribution, shown in \fref{Rtbias}A--C);
(2) the variant has a shorter mean generation interval (\fref{Rtbias}D--F); and
(3) the variant has a longer mean generation interval (\fref{Rtbias}G--I).
Then, we compare the estimated ratio $\hat{\rho}(t) = \hat{\RR}_{\textrm{var}}(t)/\hat{\RR}_{\textrm{wt}}(t)$ with the true ratio $\rho = \Rv(t)/\Rw(t)$.
%% JD: How does this convoluted LaTeX not make you crazy?
In order to allow for changes in $\RR$ driven by the introduction and lifting of non-pharmaceutical interventions \citep{flaxman2020Rt}, we model $\Rw(t)$ using a step function (\fref{Rtbias}) and let $\Rv(t) = \rho \Rw(t)$.

@@

When the assumed distribution matches the true distribution, the estimated reproduction numbers match the true values (\fref{Rtbias}A); thus, their ratio remains constant (\fref{Rtbias}B), and Rhatvar=Rvar (\fref{Rtbias}B).
However, if the generation-interval distribution of the variant differs from the assumed distribution, the ratio changes over time (\fref{Rtbias}E,H).
In particular, if the true generation intervals have a shorter mean than the assumed distribution, we over-estimate $\Rv(t)$ during the growth phase and under-estimate during the decay phase (and conversely, \fref{Rtbias}D,G), which further translates to biases in the estimated relative strength (\fref{Rtbias}E,H).
\jd{Skip this:} Therefore, biases in the estimated relative strength will correlate with the strength of the variant, (i.e., whether $\Rv(t)$ is below or above 1).

In practice, estimates of instantaneous reproduction numbers $\RR(t)$ (and therefore, their ratios) can be noisy due to limited data availability or model assumptions;
instead, we might want to estimate a single value of relative strength $\rho$.
For example, we can estimate $\rho$ by plotting the estimated strength of the variant $\hat{\RR}_{\textrm{var}}(t)$ against the estimated strength of the wild type $\hat{\RR}_{\textrm{wt}}(t)$---as presented in Figure 2 of \cite{volz2021transmission}---and performing a linear regression (\fref{Rtbias}C,F,I).
If $\rho$ is constant, and generation-interval distributions are correctly specified, we obtain a straight line with a slope of $\rho$ and intercept at zero (\fref{Rtbias}C).
However, when the assumed mean generation interval is longer than the that of the variant, we over-estimate the slope (and vice versa, \fref{Rtbias}F,I).
\jd{intercept stuff is long and vague; is that because we're not confident of the pattern?}
In theory, we expect the regression line to go through the origin, but when the assumed generation-interval distribution differs from the true distribution, this is not necessarily the case: biases in slopes due to mis-specified generation-interval distributions also lead to biases in intercepts.
This analysis suggests another method for assessing potential biases in the estimates of relative strength.

@@

We consider two types of intervention:
an intervention of constant strength, which reduces transmission by a constant factor $\theta$ regardless of age of infection ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$); and an intervention of constant speed, which reduces transmission after infection by a constant rate $\phi$ ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$).
In this case, we can control the spread of the variant when $\theta > \Rv$ or $\phi > \rv$, respectively. [[Cite speedstrength]]
We consider constant-strength and speed interventions that reduce $\Rv$ to 0.9 when both the variant and the wild type have equal mean generation intervals (\fref{strengthspeed}A--B).
While both interventions are equally effective on the strength scale (that is, $\Ry{post}=\int  K_{\mathrm{post}}(\tau) \dtau = 0.9$), they have different dynamical implications.
The constant-strength intervention affects transmission equally throughout the course of infection, whereas the constant-speed intervention has greater impact on transmission that occurs later in infection;
as a result, the constant-speed intervention reduces the post-intervention mean generation interval (\fref{strengthspeed}B) and leads to (slightly) faster exponential decay (therefore, lower $\ry{post}$).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{control.pdf}
\caption{
\textbf{Effects of constant-strength and constant-speed interventions on the spread of the new variant.}
\jd{B legend is wrong (-strength)}
(A--B) Pre-intervention (black) and post-intervention (colored) kernel of the new variant under constant-strength (A) and -speed (B) interventions assuming that the variant has equal generation intervals as the wild type.
(C--D) Pre-intervention (black) and post-intervention kernel of the new variant assuming that the variant has longer generation intervals than the wild type.
(E) Pre-intervention (black) and post-intervention strength conditional on the mean generation interval of the variant (colored).
(F) Pre-intervention (black) and post-intervention speed conditional on the mean generation interval of the variant (colored).
\jd{Something is off here; we can't have both $r=0$.} Strength and speed are calculated assuming $\rw=0$, $\delta=0.1\pday$, $\rv=0$, $\Gy{wt}=5\,\textrm{days}$, and $\kappa=1/5$ for pre-intervention conditions.
Intervention strength and speed are chosen so that post intervention strength of the new variant is 0.9 when its mean generation interval is 5 days.9.
}
\label{fig:strengthspeed}
\end{figure}

@@

The constant post-intervention speed can be understood by considering a speed-based decomposition of the kernel: $K_{\mathrm{pre}}(\tau) = \exp(\rv) b(\tau)$, where $b(\tau)$ is the initial backward generation-interval distribution and therefore integrates to 1 \citep{doi:10.1098/rspb.2020.1556}.
This decomposition allows us to re-write the post-intervention kernel (under the constant-speed intervention) as follows:
\begin{equation}
K_{\mathrm{post}}(\tau) = \exp(\rv-\phi) b(\tau).
\end{equation}
Since the post-intervention speed $\ry{post}$ is determined by the following implicit equation,
\begin{equation}
1 = \int_0^\infty \exp(-\ry{post} \tau) K_{\mathrm{post}}(\tau) \dtau = \int_0^\infty \exp(\rv-\phi-\ry{post}) b(\tau)  \dtau,
\end{equation}
we have $\ry{post} = \phi - \rv$. In particular, if speed of intervention is faster than the observed speed of spread (i.e., if $\phi > \rv$), we can control the epidemic regardless of the underlying generation-interval distribution.

@@

We explored how the generation-interval distribution shapes the link between relative strength and speed of an invadiang disease variant.
Longer generation intervals generally lead to higher relative strength for a given relative speed (and conversely, lower relative speed for a given relative strength); these can be reversed when incidence is decreasing.
Neglecting potential differences in the mean generation intervals between the variant and the wild type can bias estimates of the relative strength from incidence data;
these biases may be assessed by considering whether relative strength appears to vary systematically with underlying epidemiological dynamics.
Finally, differences in generation intervals can also lead to different conclusions about the effectiveness of interventions:
if the variant has longer (respectively, shorter) generation intervals than the wild type, speed-like (strength-like) interventions, will be more relatively more effective than naive estimates would suggest.

As the new variants, in particular the B.1.1.7 variant, of SARS-CoV-2 are spreading and becoming dominant in many countries, it is clear that the variants are more transmissible (have higher strength) than the wild type \citep{switzerland2021variant, davies2021estimated, di2021impact, leung2021early, volz2021transmission,zhao2021}.
Our analysis supports estimates of a higher strength of the new variant across a wide range of assumptions about epidemiological conditions and generation-interval distributions.
However, our analysis also shows that uncertainty in generation-interval distributions must be taken into account to obtain accurate estimates of the relative strength of variants.
If the variant has a longer infectious period \citep{kissler2021densely}, and therefore extended generation-interval distributions, current estimates may be underestimating the true relative strength.

There is currently a (minor) discrepancy in the estimates of relative strength of new variants, particularly for B.1.1.7.
Mathematical analyses have typically reported greater than 1.4 fold increase in reproduction number for the B.1.1.7 variant ($\rho > 1.4$) whereas an independent analysis of secondary attack rates from contact tracing data suggests 1.25--1.4 fold increase \citep{ukinvest}.
We can ask whether different estimates can be reconciled by generation intervals alone---for simplicity, we will assume that the variant and the wild type have identical generation-interval distributions as other studies have.

For example, \cite{davies2021estimated} estimated $\rho=1.77$ assuming a fixed generation generation interval at 5.5 days from $\delta=0.104\pday$ in their regression analysis;
similarly, \cite{volz2021transmission} estimated $\rho=1.79$ by directly comparing the ratio between $\Rv$ and $\Rw$ assuming a mean generation interval of 6.5 days and a squared coefficient of variation of 0.38.
In this case, the former study's longer mean generation interval (which typically increases $\rho$) is counter-balanced by the effects of their wider distribution.
\jd{Not clear how to interpret this. Can we match both results under a gamma assumption changing only kappa and Gbar?}
In fact, we obtain $\rho=1.81$ from $\delta=0.104\pday$, $\rw=0$, $\Gv=\Gw=6.4\days$, and $\kappa=0.38$ under the gamma assumption, explaining the lack of differences in the estimates of $\rho$ despite the differences in the assumed generation-interval distributions.

However, the fixed generation generation interval assumed by \cite{davies2021estimated} is too narrow.
The mean generation interval of 6.5 days assumed by \cite{volz2021transmission} is also likely too long:
first, this estimate actually corresponds to the observed mean serial intervals from China in between January and February, 2020 \citep{bi2020epidemiology}, during which they were more likely to observe longer serial intervals due to initial epidemic growth \citep{park2021forward, ali2020serial}; and second, recently infected individuals may be less likely to transmit after symptom onset due to increased awareness of the pandemic compared to earlier on, leading to shorter generation intervals.
If we replace the assumed generation-interval distributions for \cite{volz2021transmission} with shorter generation interval estimates from Tianjin, China ($\Gw=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.27$, which is consistent with the attack rate analysis \citep{ukinvest}.
We do not claim that these estimates of $\rho$ or $\Gw$ are more accurate;
instead, this back-of-the-envelope calculation highlights the importance of propagating uncertainty in generation-interval distributions in assessing the relative strength and speed of SARS-CoV-2 variants.
\jd{I want to drop this here. Things are getting tangled, and this \P\ is about the equal-GI case.} Obtaining accurate estimates of $\rho$ will need to take into account changes in $\rho$ as well as underlying $\rw$.

@@

We also assumed that the incidence of infection caused by the variant and the wild type is known.
However, estimating $\RR(t)$ from real data, such as case incidence, is often associated with a variety of practical challenges, including delays between infection and reporting and changes in testing patterns \citep{gostic2020practical}.
We have chosen to focus here on the underlying dynamical mechanisms that may impact inference.
Future studies may consider the development of methods based on mechanism outlined here to assess potential biases in their estimates of relative strength.

In moving from mechanism to improved estimates, we argue that both perspectives (i.e., whether we hold relative strength or speed constant) are useful in understanding the dynamics of the new variant.
Early in the spread of an emerging variant, it is likely more convenient to fix the relative speed, given that speed can be directly observed.
As more information about the transmission and immunity profiles of the new variant becomes available, we advise fixing the relative strength and inferring the speed as this assumption better matches biological mechanisms for its higher strength.
Researchers should be mindful about which quantity they hold constant and how their conclusions follow from their assumptions \citep{doi:10.1098/rspb.2020.1556}.

This study has some practical implications for analyzing the epidemiological dynamics of new variants.
First, models that assume a constant relative speed, such as standard logistic growth model, should be avoided for characterizing the growth of new variants---the relative speed changes over time depending on epidemiological conditions. \jd{need to reconcile this with advice above about Early.}
Second, the absolute strength and speed should not be neglected in favor or relative values, and both relative and absolute values should be discussed together.
While the relative strength and speed are useful for describing the spread of new variants, the absolute values determine their spread and control.
\jd{I changed this s., but maybe we can delete it; it's covered better in the next \P.} In addition, as we showed in our analysis, looking at absolute and relative values together can be critical to understanding dynamics. 
Finally, uncertainty in generation intervals should be carefully considered.

Even though SARS-CoV-2 has been spreading for more than a year, there is still considerable uncertainty in its generation intervals.
Observational studies typically focus on serial intervals (i.e., time between symptom onset of the infector and the infectee; \cite{svensson2007note}) because they are easier to measure \citep{griffin2020rapid}.
However, realized serial intervals are subject to dynamical biases that can be difficult to tease apart \citep{park2021forward}.
A few studies have tried to estimate the generation-interval distribution from serial intervals, with means ranging between 3--6 days and squared coefficients of variation ranging between 0.1--1 \citep{ferretti2020quantifying,Ferretti2020timing,ganyani2020estimating,knight2020estimating}; 
as we have shown, this uncertainty alone can generate large uncertainty in the inference of relative strength.
Furthermore, most estimates neglect dynamical biases in serial intervals caused by underlying epidemiological dynamics in their estimates of generation intervals.
Serial intervals also do not account for asymptomatic transmission, adding further uncertainty to inferences of speed and strength \citep{park2020time}.

\jd{If this \P\ is going to be mostly about us, we should say so more clearly.} Since the beginning of the SARS-CoV-2 pandemic, a few studies have emphasized the relevance of leveraging generation interval distributions to improve estimates of strength (e.g., \cite{doi:10.1098/rsif.2020.0144,park2021forward}) and the relative importance of modes of transmission (e.g., asymptomatic vs. symptomatic, \cite{park2020time}).
These same lessons should be considered when assessing the spread of variants in a partially susceptible population and improving efforts to control spread.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions.

The spread of new SARS-CoV-2 variants and the replacement of previously dominant lineages represent ongoing challenges for controlling the SARS-CoV-2 pandemic.  
By explicitly considering epidemiological context and generation interval differences together, we have shown that improving estimates of the the relative duration of infectiousness at the individual scale may represent a pathway towards more effective interventions. 
Specifically, we have shown that speed-like interventions, such as contact tracing, will be relatively more effective if variants have longer generation intervals.
Most intervention strategies throughout the current pandemic have focused on strength-like interventions \citep{flaxman2020Rt}, such as lock-downs, partly because pre-symptomatic transmission of SARS-CoV2 have limited the effectiveness of contact tracing efforts \citep{hellewell2020feasibility}.
However, given the possibility that new variants can have different infection characteristics, future studies should consider whether their transmission dynamics also differ (e.g., the amount of pre-symptomatic transmission) and evaluate intervention strategies accordingly.
